---
title: "AgeFrameworks"
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{AgeFrameworks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This Vignette introduces the tools in **BayLumPlus** that extend the original framework for age modelling. 

```{r setup}
library(BayLumPlus)
```

## Age Model Function

`Compute_AgeS_D` computes only the model age and follows the same configuration as `AgeS_Computation`.  
The only difference lies in the `DATA` parameter, which refers to a list of data inputs:

- Equivalent Doses: $D$
- Dose rate: $\dot{d}$
- Standard Error for $D$: $s_D$

Note that several prior distributions can be tested:

- **Under strict order:**
  - Nicholls
  - Approached Jeffreys:
    - *Conditional* (using successive conditional processing)
    - *StrictOrder* (using Uniform Order, UO)
  - BayLum (the current BayLum implementation on CRAN)

- **Without order:**
  - Independence

You can use the `extract_Jags_model()` function to check the list of all available models and data structures in **BayLumPlus**.

To illustrate usage, the following code snippet uses the OSL dataset from Jingbian, China  
(see the package documentation for more details).
```{r}
data("OSLJingbian")

names(OSLJingbian)
```

To avoid redundancy, we have already computed the equivalent dose model for these samples,  
as our main objective here is to illustrate the age modeling framework.


We can directly use the output of our unconstrained model, stored in `Output`, with the following line:

```{r eval=FALSE, include=FALSE}
JingbianUnconstrained = Compute_AgeS_D(list(D = OSLJingbian$D, sD = OSLJingbian$sD, ddot = OSLJingbian$sd),
                                       Nb_sample = OSLJingbian$Nb_Sample, SampleNames = OSLJingbian$SampleNames,
                                       ThetaMatrix = OSLJingbian$ThetaMatrix, prior = "Independance",
                                       PriorAge = rep(c(1, 1400), OSLJingbian$Nb_Sample),
                                       Iter = 2000, burnin = 50000, t = 10
                                       )
```

We use the *Independence* prior since we require an unconstrained Bayesian model.

To visualize the posterior marginals, we use the `plot_Ages()` function with the `"density"` mode:

```{r  fig.height=6, fig.width=7}
DoseDataset = plot_Ages(OSLJingbian$Output, plot_mode = "density")
```

## Isotonic Regression Computation

The functions `CurveIsotonic()` and `PlotIsotonicCurve()` compute the isotonic distortion of a posterior distribution 
obtained from the `Compute_AgeS_D()` function.

The currently supported prior for this isotonic correction is the *Independence* prior.

These functions require the following arguments:

- `StratiConstraints`: The stratigraphic constraint matrix. This can either be provided directly or loaded from a file.  
  If no matrix is supplied, the model assumes a strict chronological order.
  
- `Object`: The output object returned by the age model function `Compute_AgeS_D()`.

- `level`: Confidence level for High Posterior Density (HPD) regions. The default is `0.95`.

- `interactive`: An optional argument. If not provided, the program will decide whether to display an interactive  
  Directed Acyclic Graph (DAG) visualization in a web browser.


```{r Isotonic Distorsion, fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
JingbIso = PlotIsotonicCurve(StratiConstraints = c(), object = OSLJingbian$Output, level = .68)
```

The return value is a data frame containing information about the distorted posterior distribution, including:

- High Posterior Density (HPD) regions  
- Mean Bayesian estimates  
- Standard deviation estimates  

In addition to the data frame, the function also generates a visualization that displays:

- The stratigraphic constraints as a Directed Acyclic Graph (DAG)  
- A ribbon plot showing the HPD regions of the distorted posterior distribution  

Several optimization approaches are available, depending on the structure of the constraints:

- **PAVA** (Pool Adjacent Violators Algorithm) for strict chronological order  
- **Sequential Block Merging (SBM)** algorithm  
- **Global convex optimization** using the `CVXR` package  


## HPD Regions Plotting

We can compare multiple priors using the `plotHpd()` function. This function requires two main arguments:

- `listObject`: A list of output objects returned by `Compute_AgeS_D()` using different prior settings.  
- `Names`: A character vector specifying the names for each age computation method, used to distinguish them in the plot.


```{r include=FALSE}
JingbianUO = Compute_AgeS_D(list(D = OSLJingbian$D, sD = OSLJingbian$sD, ddot = OSLJingbian$ddot),
                                       Nb_sample = OSLJingbian$Nb_Sample, SampleNames = OSLJingbian$SampleNames,
                                       ThetaMatrix = OSLJingbian$ThetaMatrix, prior = "Conditional",
                            PriorAge = rep(c(1, 1400), OSLJingbian$Nb_Sample),
                            Iter = 2000, burnin = 50000, t = 10
)
```


```{r include=FALSE}
JingbianNicholls = Compute_AgeS_D(list(D = OSLJingbian$D, sD = OSLJingbian$sD, ddot = OSLJingbian$ddot),
                                       Nb_sample = OSLJingbian$Nb_Sample, SampleNames = OSLJingbian$SampleNames,
                                       ThetaMatrix = OSLJingbian$ThetaMatrix, prior = "StrictNicholls",
                            PriorAge = rep(c(1, 1400), OSLJingbian$Nb_Sample),
                            Iter = 2000, burnin = 50000, t = 10
)
```

```{r fig.height=7, fig.width=8}
plotHpd(list(JingbianUO, JingbIso, JingbianNicholls), c("UO", "Isotonic", "Nicholls"))

```


